---
title: "MK801 Genotype Processing"
author: "vivek"
date: "2/6/2021"
output: html_document
---
B6J=c("C57BL/6J"), B6N=c("BALB/cJ"), name1="B6J", name2="BALB")
is key
the control genotype file for each strain needs to be added to each batch.
the name of the strans need to agree within the file.
 
if using new strains, make sure these are changed.




# -------------------------------------------------
# PREPROCESSING CODE
# -------------------------------------------------

### The new transnetyx genotype files have two new columns. 
### In addition, Tom used numbers as AnimalID. This messes up the code.

This code processes the Genotype files to fix these two things.
1. It adds a `A_` to the `Sample Name`
2. I delete two new columns.

The processed data is placed in a new folder.  

** NOTE: The controls need to be adjusted for this by adding the `A_` to the control ID

```{r setup, include=FALSE}
library(data.table)
 
wd <- "/Users/vkumar/Box Sync/LAB/PSY todo/JCMS MK801 Analysis/MK801 Genotype Plate 1 analysis/33766-genotypes"
setwd(wd)
outputDir <- "../Processed"
dir.create(file.path("../Processed/"), showWarnings = FALSE)


inputfiles = list.files(wd, full.names=FALSE, pattern=".txt")

for (inputfile in inputfiles){
    tbl1 <- fread(inputfile, sep = '\t', colClasses = c("Sample Name" = "character"), skip = 0) # new on Feb 2024
    names(tbl1)<-make.names(names(tbl1),unique = TRUE)

    tbl1$Sample.Name <- paste("A", tbl1$Sample.Name, sep="_") # Added this because Tom has only numbers for Sample Name, this can be removed
    tbl1 <- tbl1[, -c(3:4)] # added in 2024 to remove two 
    
    outfile1 <- paste(outputDir, inputfile, sep = "/")
  write.table(tbl1, outfile1, row.names = FALSE, quote = FALSE, sep = "\t")
  }


```


## correct the input file name to the folder of processed files from the code above
## Asaf's code to combine the genotypes processed above
 
```{r}
 
filepath <- "/Users/vkumar/Box Sync/LAB/PSY todo/JCMS MK801 Analysis/MK801 Genotype Plate 1 analysis/Processed"

setwd(filepath)
 
inputfiles <- Sys.glob(paste0(filepath,"/*.txt"))
 
read_neogen <- function(inputfiles, outfile, B6J=c("A_020269"), B6N=c("A_018606"), name1="B6J", name2="B6N"){
  # Read Karl Broman analysis result with unique and mapped markers, save a list of the unique only
  kb <- read.csv(url("https://raw.githubusercontent.com/kbroman/MUGAarrays/master/UWisc/mini_uwisc_v2.csv"))
  kblist <- kb$marker[kb$unique & (!kb$unmapped)]
  # Read the Neogen input file and write the genotypes matrix into the output csv file
  tbl <- NULL
  for (inputfile in inputfiles){
    tbl1 <- read.delim(inputfile, skip = 1, header = F)
    tbl <- rbind(tbl, tbl1)
  }
  w3 <- reshape(tbl[,c(1:2,4)], v.names="V4", idvar = "V1", timevar="V2", direction="wide")
  rownames(w3) <- w3$V1
  w3 <- w3[,-1]
  colnames(w3) <- gsub("^V4.", "", colnames(w3))
  #w3 <- t(w3)
  w4 <- reshape(tbl[,c(1:2, 5)], v.names="V5", idvar = "V1", timevar="V2", direction="wide")
  rownames(w4) <- w4$V1
  w4 <- w4[,-1]
  colnames(w4) <- gsub("^V5.", "", colnames(w4))
  #w4 <- t(w4)
   
  # Find the B6J and B6N consensus. If one allele is different remove the marker
  # If B6J==B6N remove the marker
  keepr <- apply(w3[B6J, ,drop=F], 2, function(col) length(unique(col)) == 1) &
    apply(w4[B6J, ,drop=F], 2, function(col) length(unique(col)) == 1) &
    apply(w3[B6N, ,drop=F], 2, function(col) length(unique(col)) == 1) &
    apply(w4[B6N, ,drop=F], 2, function(col) length(unique(col)) == 1) &
    w3[B6J[1],] == w4[B6J[1],] &
    w3[B6N[1],] == w4[B6N[1],] &
    w3[B6J[1],] != w3[B6N[1],]
   
  # Set the allele names to A and B for B6J and B6N
  w3 <- w3[, keepr]
  w4 <- w4[, keepr]
  d3 <- w3 == w3[rep(B6J[1],nrow(w3)), ]
  d3[d3==TRUE] = "A"
  d3[d3==FALSE] = "B"
  d4 <- w4 == w4[rep(B6J[1],nrow(w4)), ]
  d4[d4==TRUE] = "A"
  d4[d4==FALSE] = "B"
   
  # Concatenate the two alleles
  for (i in 1:nrow(d3)) d3[i,] <- paste0(d3[i,], d4[i,])
  d3 <- d3[, intersect(colnames(d3), kblist)]
  # Separate the founders and write them in the first two rows followed by the rest
  dout <- rbind(d3[B6J[1],,drop=F], d3[B6N[1],,drop=F], d3[!rownames(d3) %in% c(B6N, B6J),])
  rownames(dout) <- c(name1, name2, rownames(dout)[3:nrow(dout)])
  write.csv(dout, file=outfile)
}
 
read_neogen(inputfiles,"new_outfile.csv")
 
```